---
title: "Notes"
author: "Andrew Abisha Hu"
date: "8/20/2018"
output: html_document
---
Start
```{Basic}
setwd("/users/andrewhu/desktop/final/en_US")
con <- file("en_US.twitter.txt", "r") 
readLines(con, 1) ## Read the first line of text 
readLines(con, 1) ## Read the next line of text 
readLines(con, 5) ## Read in the next 5 lines of text 
close(con)
```

You might want to create a separate sub-sample dataset by reading in a random subset of the original data and writing it out to a separate file. That way, you can store the sample and not have to recreate it every time. You can use the rbinom function to "flip a biased coin" to determine whether you sample a line of text or not.

Regular expression:

Leterals: the exact matching with the lines.

However, we need a way to express:

        whitespace word boundaries
        set of literals
        the begining and end of a line
        alternatives
        
So, we need metacharacters.

        
### "^" represents the start of a line

^i think

will match:

```
i think we all ....
i think i have been
i think....
```
but it will not match if the "i think"" is at the middle of the line

### $ represents the end of a line

morning$

will match:

```
..... this morning
then had to ...in the morning
```

### [] will accept at a given point

[Bb][Uu][Ss][Hh]

will match all kinds of combos of bush

---



### combos:


^[Ii] am

will macth:

I am 

i am

---

### specify a range of letters

^[0-9][a-zA-Z]

it will match for a number at the begining followed by any character

---

[^?.]$ will look for any that is **not ending** with ? and .

```
i like basketballs
6 and 9
dont worry....... we all die anyway!

```

. refers to any character

9.11

will match:

```
...... 9-11 rules
if.... 9/11....
9.11
9:11
```

| indicates the alternative (or)

flood|earthquake|
will fit with the lines that have flood or earthquake

---

^[Gg]ood|[Bb]ad

will fit the begining of Good/good/Bad/bad

note: the "bad" does not need to be at the begining


^ ([Gg]ood|[Bb]ad) then restrains that the bad should be at the begining of the line




```{r cars}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
