---
title: "Reproducible Research Project 2"
output:
  pdf_document: default
  html_document: default
---

# Synopsis

The basic goal of this assignment is to explore the NOAA Storm Database and answer two questions: which types of events are most harmful to population health and which types of events have the greatest economic consequences. From the data set, we found out that **TORNADO** has the largest impact on damaging both population and economy.

# Data Processing 

## Reading the raw data
```{r reading data ,echo=TRUE}
storm<- read.csv("/users/andrewhu/desktop/storm.csv")
```


## Previewing the structure of the data
```{r Previwing the raw data ,echo=TRUE}
head(storm)
dim(storm)
```

## Finding the variables we are interested

The columns we are intersted related to the **harmfulness of Population**, are the "Fatalities" and "Injuries". Here we take a look of their summaries:
```{r summaries for population damages ,echo=TRUE}
summary(storm$FATALITIES)
summary(storm$INJURIES)
```

Create a variable called **dmg_pop** indicating the damage of the population, combining the fatalities and injuries.
```{r dmg_pop ,echo=TRUE}
storm$dmg_pop = storm$FATALITIES + storm$INJURIES
```

Calculate the dmg_pop by each type of events and return a new data frame

```{r popdamage ,echo=TRUE}
popdamage <- aggregate(dmg_pop~EVTYPE, data=storm, sum)

```

Simply taking a look of the new data frame we just create, we found that there are a lot of EVTYPE, and many of the EVTYPE contain 0 dmg_pop

```{r ,echo=TRUE}
summary(popdamage)
head(popdamage)
```

Hence, we need to "summary" the popdamage data frame. We can subset a data frame which contains top 5 damages for each EVTYPE.
```{r ,echo=TRUE}
library(dplyr)
popdmgtop5 <- popdamage %>% arrange(desc(dmg_pop)) %>% slice(1:5)
```


Now, the variables we are interested for population damage related are processed finished. Let's take a loot at economic damage-related variables, which are "PROPDMG" and "CROPDMG".


```{r ,echo=TRUE}
summary(storm$PROPDMG)
summary(storm$CROPDMG)
```

Then, just as the steps we created for the damage for population above, we simply create a variable indicating the total impact for the economic damage, combining the two variables.

```{r ,echo=TRUE}
storm$dmg_eco <- storm$PROPDMG + storm$CROPDMG
```

Now, calculate the sum of dmg_eco for each event type and return a data frame

```{r ,echo=TRUE}
ecodamage <- aggregate(dmg_eco ~ EVTYPE, data=storm, sum)
```

Simply take a look at the new data frame for eco damage:

```{r ,echo=TRUE}
head(ecodamage)
```


Filter the ecodamage for containing top 5 damges of EVTYPE only:

```{r ,echo=TRUE}
library(dplyr)
ecodmgtop5 <- ecodamage %>% arrange(desc(dmg_eco)) %>% slice(1:5)
```

# Results

Now, simply printing out the popdmgtop5 and ecodmgtop5, we can have an idea of which EVTYPE has the largest impact on the population and economy:
```{r ,echo=TRUE}
popdmgtop5
ecodmgtop5
```

In addition, let's do some plots.

Population Damage:
```{r ggplot for popdmgtop5 ,echo=TRUE}
library(ggplot2)
##ggplot
#base
g<- ggplot(popdmgtop5, aes(x=EVTYPE, y=dmg_pop))
#
g + geom_bar(stat= "identity") + labs(title= "Top 5 Events for Population Damage")

```

Economic Damage:
```{r ggplot for ecodmgtop5 ,echo=TRUE}
##ggplot
#base
g<- ggplot(ecodmgtop5, aes(x=EVTYPE, y=dmg_eco))
#
g + geom_bar(stat= "identity") + labs(title= "Top 5 Events for Economic Damage")

```


