---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---



```{r}
#-------- Tutorial
library(data.table)

#Example: Create a blank vector, and duplicate it.

nums <- c(2,4,6,8)
dup <- numeric() #the blank one
for (num in nums) {
        dup<- c(dup,nums) #the new dup=  blank dup + nums
        
}


#part 1

pollutantmean <- function(directory, pollutant, id= 1:332) {
        
        filelist<- list.files(path=directory, pattern= ".csv", full.names=TRUE)#Read the list of files, using full route ( full.names)
        values<- numeric() #Create a numeric vector, make sure to create before the for loop
        
        for (i in id){
        data<-read.csv(filelist[i]) # read the filelist
        values <- c(values, data[[pollutant]])   # add the value into the blank vector you created before, use double [] to locate the var  
        }
        mean(values,na.rm=TRUE)
        
}

pollutantmean("/users/andrewhu/desktop/Coursera/specdata","sulfate")



```
```{r}
#part2


data<- read.csv("specdata/001.csv")
#get the number of complete cases
sum(complete.cases(data)) #117 complete cases in 001.csv


```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

