---
title: "Quiz"
output: html_document
---
Question 1
What is the variance of the distribution of the average an IID draw of nn observations from a population with mean \muμ and variance \sigma^2σ 
2

Ans: sigma^2 / n


Question 2
Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a DBP less than 70?

Ans:
```{r setup}
pnorm(70,mean=80,sd=10)
```
Q3:

Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?

Ans:
```{r }
qnorm(0.95, mean=1100, sd=75)
```
Q4: Sample distribution
Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is the 95th percentile of the distribution of that sample mean?

```{r pressure, echo=FALSE}
qnorm(0.95,mean=1100,sd=75/sqrt(100))
```

Q5: Binomial

You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?

```{r}
pbinom(3, size = 5, prob = 0.5, lower.tail = FALSE)
```
Q6: Sample Distribution

The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?

For Standard error: STD/ sqrt(n)
```{r}
10/ sqrt(100)
```

Calculating the probability between 14 and 16:

```{r}
pnorm(16, mean = 15, sd = 1) - pnorm(14, mean = 15, sd = 1)
```
Q7:

Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?

Ans: It should be near the mean, .5


Q8:Poisson


The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 or fewer people?
```{r}
ppois(10, lambda = 15)
```

In a population of interest, a sample of 9 men yielded a sample average brain volume of 1,100cc and a standard deviation of 30cc. What is a 95% Student's T confidence interval for the mean brain volume in this new population?

```{r}
#Formula : 
#ts= qt(1 - alpha / 2, n - 1)
# mean + - ts*  std/sqrt(n)

1100 + c(-1, 1) * qt(1- 0.05/2, 9-1) * 30 / sqrt(9)
```
A diet pill is given to 9 subjects over six weeks. The average difference in weight (follow up - baseline) is -2 pounds. What would the standard deviation of the difference in weight have to be for the upper endpoint of the 95% T confidence interval to touch 0?

```{r}
# mean + - ts*  std/sqrt(n)
#ts= qt(1- alpha/2, n-1)
ts = qt(0.975,8) #2.306
#s <- -x_bar*sqrt(n) / ts

-2*sqrt(9)/x
```


