---
title: "Week4 Notes"
author: "Andrew Abisha Hu"
date: "8/18/2018"
output: html_document
---

## Generating Random Numbers

Functions for probability distributions in R
- `rnorm`: generate random Normal variates with a given mean and standard deviation
- `dnorm`: evaluate the Normal probability density (with a given mean/SD) at a point (or vector of points)
- `pnorm`: evaluate the cumulative distribution function for a Normal distribution 
- `rpois`: generate random Poisson variates with a given rate

---

Probability distribution functions usually have four functions associated with them. The functions are prefixed with a
- `d` for density
- `r` for random number generation 
- `p` for cumulative distribution
- `q` for quantile function

## Simulating data in r:
```{r}
x<- rnorm(n, mean = 0, sd = 1)
x<- dnorm(x, mean = 0, sd = 1, log = FALSE)
```


## Simulating data from a linear model:
```{r}
set.seed(20)
x <- rnorm(100)
e <- rnorm(100, 0, 2)
y <- 0.5 + 2 * x + e
summary(y)
```


