---
title: "Final Project, Regression Model"
output:
  pdf_document: default
  html_document: default
---
# Executive Summary:

This project is set to explore the relationship between a set of variables and miles per gallon. We need to answer: 

1. Is an automatic or manual transmission better for MPG 

2. Quantify the MPG difference between automatic and manual transmissions


After analyzing the ```mtcars``` data, we can conclude that manual transmission produces more mpg compared to auto transmission. And according to our best fitted model, manual transmission achieve 2.936 more mpg than auto transmission


# Analysis:


Loading libraries and datasets
```{r setup}
library(datasets)
library(ggplot2)
library(dplyr)
data(mtcars)
```

Transformation:
```{r}
mtcars<- mutate(mtcars, am=factor(mtcars$am,labels=c("Auto","Manual")), vs=factor(vs),gear=factor(gear), carb=factor(carb))
```


Basic preview:
```{r cars}
head(mtcars)
```


Exploratory analysis (Boxplot for Transmission method vs. MPG)
```{r pressure}
g <- ggplot(mtcars, aes(x=factor(am), y= mpg))
g + geom_boxplot(aes(fill=factor(am))) + ggtitle("Transmission vs MPG")
```


## Building models:

Method 1: (reference: how to build nested model in R)
```{r}
fit <- lm(mpg ~ factor(am), data = mtcars)
fit2 <- update(fit, mpg ~ factor(am) + wt)
fit3 <- update(fit, mpg ~ factor(am)+ wt + hp)
fit4 <- update(fit, mpg ~ factor(am)+ wt + hp + qsec)
fit5<- update(fit, mpg ~ factor(am)+ wt + hp + qsec + cyl)
anova(fit, fit2, fit3, fit4, fit5)#use anova table to test whether you should include certain variables
```

Method 2: Stepwise

```fit_step <- step(fit, direction="both")```
```{r,include=FALSE}
fit<- lm(mpg~. ,data=mtcars)
fit_step <- step(fit, direction="both")
```


Comparison:
```{r}
summary(fit)
summary(fit_step)
```
**Conclusion** :

Comparing the initial model (putting all the variables) and our best model(fit_step), we can conclude that the original model has a 0.78 Adjusted R square, meaning that there is only 78% of the variables is explained by this model. However, we have a higher Adjusted R square, 0.834, in our best fitted model.

** Coefficient interpretation **

For every other variables stay the same, manual transmission will increase 2.936 more mpg, compared to auto transmission


## Model diagnosis

Residual analysis:
```{r}
plot(fit_step)
```

**Residuals vs. Fitted :**

The residuals are scattered, ensuring the independence between fitted values and residuals. If there is any pattern, then we should change the model.

**Q-Q Plot:**

The points are mostly closed with the line. Hence, we can suggest that the residuals are nornally distributed.


## Statistical inference:

**T-test**
```{r}
t.test(mpg~am, mtcars)
```

**Conclusion**

According to the t-test, we can reject the null hypothesis that the transmission method will not have an impact on the mpg.
