---
title: "Quiz"
output: html_document
---
Consider the data set given below
x <- c(0.18, -1.54, 0.42, 0.95)
And weights given by
w <- c(2, 1, 3, 1)

Give the value of \mu that minimizes the least squares equation

Sigma = Wi (Xi - µ) ^2


```{r, include=FALSE}
x <- c(0.18, -1.54, 0.42, 0.95)
w <- c(2, 1, 3, 1)

sum(x*w) / sum(w)


```

Consider the following data set

x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

Fit the regression ** through the origin **  and get the slope treating y as the outcome and x as the regressor.
(Hint, do not center the data since we want regression through the origin, not through the means of the data.)


```{r }
#no intercept
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

fit<- lm(y~x -1)
fit$coefficients
```

Do data(mtcars) from the datasets package and fit the regression model with mpg as the outcome and weight as the predictor. Give the slope coefficient.

```{r pressure}
data(mtcars)
fit <- lm(mpg~wt, data= mtcars)
fit$coefficients
```

Consider the following data set (used above as well). What is the intercept for fitting the model with x as the predictor and y as the outcome?

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

fit<- lm(y~x)
fit$coefficients

```

Consider the data given by
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
What value minimizes the sum of the squared distances between these points and itself?

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
mean(x)
```


Let the slope having fit Y as the outcome and X as the predictor be denoted as β1. Let the slope from fitting X as the outcome and Y as the predictor be denoted as γ1. Suppose that you divide β1 by γ1; in other words consider β1/γ1. What is this ratio always equal to?


```{r}
cor(X, Y)*sd(Y)/sd(X) / (cor(X,Y)*sd(X)/sd(Y))
= sd(Y)^2/(sd(X)^2) 
= var(Y)/var(X)
```

Consider the following data with x as the predictor and y as as the outcome.

x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)

Give a P-value for the two sided hypothesis test of whether \beta_1 β 
from a linear regression model is 0 or not.

```{r}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
fit= lm(y~x)
summary(fit)
```

In the \verb|mtcars|mtcars data set, fit a linear regression model of weight (predictor) on mpg (outcome). Get a 95% confidence interval for the expected mpg at the average weight. What is the lower endpoint?

```{r}
data(mtcars)
y <- mtcars$mpg
x <- mtcars$wt
fit = lm(y~x, data= mtcars)

predict(fit, newdata= data.frame(x= mean(x)),interval= "confidence")
```

Consider again the \verb|mtcars|mtcars data set and a linear regression model with mpg as predicted by weight (1,000 lbs). A new car is coming weighing 3000 pounds. Construct a 95% prediction interval for its mpg. What is the upper endpoint?

```{r}
predict(fit, newdata = data.frame(wt = 3), interval = ("prediction"))
```


Consider again the \verb|mtcars|mtcars data set and a linear regression model with mpg as predicted by weight (in 1,000 lbs). A “short” ton is defined as 2,000 lbs. Construct a 95% confidence interval for the expected change in mpg per 1 short ton increase in weight. Give the lower endpoint.

```{r}
fit2 = lm(mpg~I(wt/2), mtcars)
confint(fit2,level=0.95)

```


Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.


```{r}
library(datasets)
data(mtcars)

fit <- lm(mpg~factor(cyl)+ wt, data=mtcars)
fit
```

Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models. Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?.

```{r}
fit<- lm(mpg~factor(cyl)+wt,data=mtcars)
fit2 <- lm(mpg~factor(cyl),data=mtcars)
fit
fit2

```


Consider the mtcars data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.


```{r}
fit<- lm(mpg~factor(cyl)+wt,mtcars)
fit2<- lm(mpg~factor(cyl)*wt, mtcars)
anova(fit,fit2,test="Chisq")
```


Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight inlcuded in the model as

lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)

How is the wt coefficient interpretted?


```{r}
summary(lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars))
```



Consider the following data set

x <- c(0.586, 0.166, -0.042, -0.614, 11.72)

y <- c(0.549, -0.026, -0.127, -0.751, 1.344)

Give the hat diagonal for the most influential point:
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
fit<- lm(y~x)
hatvalues((fit))
```



Consider the following data set

x <- c(0.586, 0.166, -0.042, -0.614, 11.72)

y <- c(0.549, -0.026, -0.127, -0.751, 1.344)

Give the slope dfbeta for the point with the highest hat value.


```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
fit<- lm(y~x)
dfbetas(fit)[,2]
```


Consider the space shuttle data \verb|?shuttle|?shuttle in the \verb|MASS|MASS library. Consider modeling the use of the autolander as the outcome (variable name \verb|use|use). Fit a logistic regression model with autolander (variable auto) use (labeled as "auto" 1) versus not (0) as predicted by wind sign (variable wind). Give the estimated odds ratio for autolander use comparing head winds, labeled as "head" in the variable headwind (numerator) to tail winds (denominator).

```{r}
library(MASS)
data("shuttle")
summary(shuttle)

#Create a binary variable
shuttle <- mutate(shuttle, auto=1* (use=="auto"), wind=factor(wind))

#glm model
fit1 <- glm(auto~wind-1,family = binomial(link="logit"), shuttle)
summary(fit1)

#Calculate Odds Ratio
coe <- summary(fit1)$coef

#Ans
exp(coe[1,1]-coe[2,1])
```


Consider the previous problem. Give the estimated odds ratio for autolander use comparing head winds (numerator) to tail winds (denominator) adjusting for wind strength from the variable magn.

```{r}
fit2 <- glm(auto~wind + factor(magn)-1, family=binomial(link="logit"), shuttle)
coe2 <-summary(fit2)$coef
coe2
exp(coe2[1,1]-coe2[2,1])

```


