---
title: "Week1"
output: html_document
---

##  Regression to the mean:

* Suppose that we normalize X (child's height) and Y (parent's height) so that they both have mean 0 and variance 1. 
* Then, recall, our regression line passes through (0, 0), (the mean of the X and Y).
* If the slope of the regression line is Cor(Y,X), regardless of which variable is the outcome (recall, both standard deviations are 1).
* Notice if X is the outcome and you create a plot where $X$ is the horizontal axis, the slope of the least squares line that you plot is 1/Cor(Y, X). 

```{r }
library(UsingR)
data(father.son)
#Normalize
y <- (father.son$sheight - mean(father.son$sheight)) / sd(father.son$sheight)
x <- (father.son$fheight - mean(father.son$fheight)) / sd(father.son$fheight)

#correlation
rho <- cor(x, y)

```


## Fitting the data: Diamond exercise
```{r cars}
library(UsingR)
data(diamond)
fit <- lm(price ~ carat, data = diamond)
coef(fit)
# For every increase one unit increase in carat, we expect the price will increase by 3721.0249
```

## Getting a more interpretable intercept
```{r, echo = TRUE}
#mean center the predictor variables, carat
fit2 <- lm(price ~ I(carat - mean(carat)), data = diamond)
coef(fit2)
# Intercept: Expected price for the average sized diamond, in this case is 0.2042
```

## Predicting the price of the diamonds
```{r pressure, echo=FALSE}
newx <- c(0.16, 0.27, 0.34)
predict(fit, newdata = data.frame(carat = newx))
```

## Residuals

Represents variation left unexplained by our model.




