---
title: "R-Programming Notes"
author: "Andrew Abisha Hu"
date: "8/18/2018"
output: html_document
---
## **Table of content**

* **Week1**

        * Assign values
        
        * Subset with conditions
        
        * Check missing values

* **Week2**

        * Write an if-else expression

        * Write a for loop, a while loop, and a repeat loop

        * Define a function in R and specify its return value [see Functions Part 1 and Part 2]

        * Describe how R binds a value to a symbol via the search list

        * Define what lexical scoping is with respect to how the value of free variables are resolved in R

        * Describe the difference between lexical scoping and dynamic scoping rules

        * Convert a character string representing a date/time into an R datetime object. [see Dates and Times]

* **Week3** 

        * apply
        
        * tapply
        
        * sapply
        
        * lapply
        
        * mapply
        
* **Week4**

        * Generate random numbers
        
---

## Reading url:
```{r }
Con <-url("http://www.jhsph.edu" ,"r" )
Con
x<- readLines(Con)
head(x)
```

## Subsetting:
```{r }
# double braket can be used for computed index

x<- list(foo=1:4, bar=0.6, baz= "hello")
x$bar
x$foo
name <- "foo"
x[[name]]

x<- list(a= list(10,12,14), b=c(3.14,2.81))
x[[c(1,3)]]

x[[c(2,1)]]

x<- matrix(1:6, 2,3 )

x[1,2]
x[1,]
x[,2]

#Return the matrix
x[1,2,drop=F]
x[2, ,drop=F]

#Partial matching
x<- list(adjdofjodaif= 1:5)
x$a
x[["a"]]
x[["a", exact=F]]

#Removing NA values
x<- c(1,2,3,NA,NA,5)
bad<- is.na(x) #create a logical vector
x[!bad]

#choosing non NA values
y<- c("a", "b","c", NA, NA,"d")

good<- complete.cases(x,y)
x[good]
y[good]
```


## Logical:
```{r}
#Double equal sign tests the equality 
x <- c(17, 14, 4, 5, 13, 12, 10)
x[x > 10]<- 4
x[x == 10] <-4
x
#Multiple conditions subsetting
ex<- hw1[hw1$Ozone>31 & hw1$Temp>90, ]
summary(hw1[hw1$Month==5,])
```

## **For-loops:**
```{r}
x<- c("a","b","c","d")

for (i in 1:4) { print(x[i])}

for ( i in seq_along(x)) { print(x[i])}

for (letter in x) {print(letter)}

# Nested for loops 

x<- matrix(1:6, 2, 3)

for( i in seq_len(nrow(x))) { 
        
        for (j in seq_len(ncol(x)))
        { print(x[i,j])
                
        }
}
```

## **First function practice**
```{r}
#First function

above10 <- function(x){
        use<- x>10 
        x[use]
        
}

above<- function(x,n=10){
        use<- x>n
        x[use]
        
}

#Matrix column mean
columnmean<- function(y, removeNA= TRUE) {
        nc<- ncol(y)
        means<- numeric(nc)
        for (i in 1:nc) { 
                means[i]<- mean(y[,i], na.rm=removeNA )}
        means
        
}

#Functions Part I

#Argument Matching

#The "... argument in the functions indicates the pass along to other functions. will address more in the future


```

## Lexical Scoping
```{r lexical}
make.power<- function(n){
        pow <- function(x) {
                x^n
        }
        pow
}

cube<- make.power(3)#only specify n, does not specify x.
cube 
cube(3)# the 3 here is specifying x

square<- make.power(2)
square(3)

#Looking for functions' environment

ls(environment(cube))
get("n",environment(cube))

#Dates
x<- as.Date("2012-01-01")
y<- strptime("9 Jan 2011 11:34:21", "%d %b %Y %M:%S")
x
y
x-y

x<- as.POSIXlt(x)
x
x-y
```

## **Swirl Practice: Logical operator**
```{r}
# & evaluates all
# && evaluate the first member of a vector (operand)


# | is "or" , evaluates the whole vector (operand)
# same logic, || only evaluates the first member of a right operand
TRUE | c(TRUE, FALSE, FALSE)


5 > 8 || 6 != 8 && 4 > 3.9
#First, the && is evaluated, is true
#Then, 5<8 is false, so now is FALSE || TRUE
# || only evaluates the first element of the right operand, hence, it is true.


#isTRUE
isTRUE(6>4)

#identical()

#xor() : Stands for the exclusive "OR"
#If there is a or, then it will be TRUE.
#If there is not a or (two arguments are the same), then will be FALSE

xor( 5==6, !FALSE) # (FALSE,TRUE) ==> TRUE

#which() : return the location of the true value 
ints <- sample(10)
which(ints>7)

#any() : return Ture if one or more of the elements are true
#all(): if all of the elements are true
```



**Swirl practice: Functions**
```{r}

# %% : Return the remainder
remainder <- function(num, divisor=2) {
 num %% divisor
        
}

remainder(5)
# See function's arguments:

args(remainder)

#Paste: concatenate

telegram <- function(...){
        paste("START",...,"STOP")
  
}
```


**Create Matrix**
```{r}

#Create Matrix

my_matrix2<- matrix(1:20,nrow=4, ncol=5)
my_matrix<- matrix(1:9,byrow=TRUE,ncol=5) #byrow= filled by the rows

#Cbind: combine column ( add a column)

patients<- c("Bill","Gina","Kelly","Sean")
cbind(patients,my_matrix2) #Note: the numbers will become "" because matrix can only have one type of data

```

**Create data frame**
```{r}
#Data frame: can have different kinds of data

my_data<- data.frame(patients,my_matrix2)

#Create names for columns

cnames<-c("patient","age","weight","bp","rating","test")

colnames(my_data)<-cnames

#Same, use rownames to create for rows
```

---

## **Apply**

`apply` is used to a evaluate a function (often an anonymous one) over the margins of an array.

- It is most often used to apply a function to the rows or columns of a matrix

- It can be used with general arrays, e.g. taking the average of an array of matrices 

```r
> str(apply)
function (X, MARGIN, FUN, ...)
```

- `X` is an array
- `MARGIN` is an integer vector indicating which margins should be “retained”. 
- `FUN` is a function to be applied
- ... is for other arguments to be passed to `FUN`

- `rowSums` = `apply(x, 1, sum)`
- `rowMeans` = `apply(x, 1, mean)`
- `colSums` = `apply(x, 2, sum)`
- `colMeans` = `apply(x, 2, mean)`

## Looping on the Command Line

Writing for, while loops is useful when programming but not particularly easy when working interactively on the command line. There are some functions which implement looping to make life easier.

- `lapply`: Loop over a list and evaluate a function on each element 

- `sapply`: Same as `lapply` but try to simplify the result

- `apply`: Apply a function over the margins of an array

- `tapply`: Apply a function over subsets of a vector

- `mapply`: Multivariate version of `lapply`

An auxiliary function `split` is also useful, particularly in conjunction with `lapply`.

```{r }
#Apply (Margin 1 indicates row, 2 is cols)
library(datasets)
apply(iris[, 1:4], 2, mean)

#mapply: apply functions to multiple vector
str(mapply)

#tapply: apply functions to subset of a vector
str(tapply) #index is a factor, or a list of factor
data("mtcars")

#How can one calculate the average miles per gallon (mpg) by number of cylinders in the car (cyl)?

tapply(mtcars$mpg,mtcars$cyl,mean)

#split takes a vector or other objects and splits it into groups
str(split) #x= vector or df, f= factor or list of factor, drop= empty factors levels should be dropped?

x<- c(rnorm(10),runif(10),rnorm(10,1))
f<- gl(3,10)
split(x,f)

#combine with lapply (the same with using just tapply)
lapply(split(x,f),mean)
tapply(x,f,mean)

#examples
library(datasets)
head(airquality)

#split the data into monthly basis

s<- split(airquality,airquality$Month)

#applying with anonymous functions
lapply(s,function(x) colMeans(x[, c("Ozone","Solar.R","Wind")]))

sapply(s,function(x) colMeans(x[,c("Ozone","Solar.R","Wind")]))

sapply(s,function(x) colMeans(x[,c("Ozone","Solar.R","Wind")],na.rm=TRUE))
```


---

## Generating Random Numbers

Functions for probability distributions in R
- `rnorm`: generate random Normal variates with a given mean and standard deviation
- `dnorm`: evaluate the Normal probability density (with a given mean/SD) at a point (or vector of points)
- `pnorm`: evaluate the cumulative distribution function for a Normal distribution 
- `rpois`: generate random Poisson variates with a given rate

---

Probability distribution functions usually have four functions associated with them. The functions are prefixed with a
- `d` for density
- `r` for random number generation 
- `p` for cumulative distribution
- `q` for quantile function

## Simulating data in r:
```{r}
x<- rnorm(n, mean = 0, sd = 1)
x<- dnorm(x, mean = 0, sd = 1, log = FALSE)
```


## Simulating data from a linear model:
```{r}
set.seed(20)
x <- rnorm(100)
e <- rnorm(100, 0, 2)
y <- 0.5 + 2 * x + e
summary(y)
```

