---
title: "Project 2"
author: "Andrew Abisha Hu"
date: "8/16/2018"
output: html_document
---

## **Instructions:**

The Hospital Compare web site contains a lot of data and we will only look at a small subset for this
assignment. The zip file for this assignment contains three files

• outcome-of-care-measures.csv: Contains information about 30-day mortality and readmission rates

for heart attacks, heart failure, and pneumonia for over 4,000 hospitals.
• hospital-data.csv: Contains information about each hospital.

• Hospital_Revised_Flatfiles.pdf: Descriptions of the variables in each file (i.e the code book).

A description of the variables in each of the files is in the included PDF file named Hospital_Revised_Flatfiles.pdf.

This document contains information about many other files that are not included with this programming assignment. You will want to focus on the variables for Number 19 (\Outcome of Care Measures.csv") and
Number 11 (\Hospital Data.csv"). You may find it useful to print out this document (at least the pages for Tables 19 and 11) to have next to you while you work on this assignment. In particular, the numbers of the variables for each table indicate column indices in each table (i.e. \Hospital Name" is column 2 in the outcome-of-care-measures.csv file).

## Part I

Write a function called best that take two arguments: the 2-character abbreviated name of a state and an
outcome name. The function reads the outcome-of-care-measures.csv file and returns a character vector
with the name of the hospital that has the best (i.e. lowest) 30-day mortality for the specified outcome
in that state. The hospital name is the name provided in the Hospital.Name variable. The outcomes can
be one of \heart attack", \heart failure", or \pneumonia". Hospitals that do not have data on a particular
outcome should be excluded from the set of hospitals when deciding the rankings.
Handling ties. If there is a tie for the best hospital for a given outcome, then the hospital names should
be sorted in alphabetical order and the first hospital in that set should be chosen (i.e. if hospitals \b", \c",
and \f" are tied for best, then hospital \b" should be returned).


The function should use the following template.

```
best <- function(state, outcome) {
## Read outcome data
## Check that state and outcome are valid
## Return hospital name in that state with lowest 30-day death
## rate
}

```
The function should check the validity of its arguments. If an invalid state value is passed to best, the
function should throw an error via the stop function with the exact message \invalid state". If an invalid
outcome value is passed to best, the function should throw an error via the stop function with the exact
message \invalid outcome".


```{r setup}
library(dplyr)
best <- function(state, outcome) {
        ## Read outcome data
        data <- read.csv("outcome-of-care-measures.csv",colClasses = "character") 
        
        data<- select(data,c(2,7,11,17,23)) #subset for hospital, state, heart attack, heart failure and pneumonia
        colnames(data) <- c("hospital", "state", "heart attack", "heart failure", "pneumonia")
        
        # Check that state and outcome are valid
        if (!state %in% data[ ,"state"]){ stop('invalid state')}
        else if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) {stop('invalid outcome')} #for invalid
        else {
        state_logical <- which(data[, "state"] == state)
        df_state <- data[state_logical, ]    # subset data for having state
        ot <- as.numeric(df_state[, eval(outcome)])#turn the outcome into numeric
        min_eval <- min(ot, na.rm = TRUE) #assign the min value
        result  <- df_state[, "hospital"][which(ot == min_eval)]
        output  <- result[order(result)] }
return(output)     
}
best("SC", "heart attack") 
```

## **Part II**

Write a function called rankhospital that takes three arguments: the 2-character abbreviated name of a
state (state), an outcome (outcome), and the ranking of a hospital in that state for that outcome (num).
The function reads the outcome-of-care-measures.csv file and returns a character vector with the name
of the hospital that has the ranking specified by the num argument. For example, the call
rankhospital("MD", "heart failure", 5)
would return a character vector containing the name of the hospital with the 5th lowest 30-day death rate
for heart failure. The num argument can take values \best", \worst", or an integer indicating the ranking
(smaller numbers are better). If the number given by num is larger than the number of hospitals in that
state, then the function should return NA. Hospitals that do not have data on a particular outcome should
be excluded from the set of hospitals when deciding the rankings.
Handling ties. It may occur that multiple hospitals have the same 30-day mortality rate for a given cause
of death. In those cases ties should be broken by using the hospital name.

```{r cars}
library(dplyr)
rankhospital <- function(state, outcome, rank) {
        ## Read outcome data
        data <- read.csv("outcome-of-care-measures.csv",colClasses = "character") 
        
        data<- select(data,c(2,7,11,17,23)) #subset for hospital, state, heart attack, heart failure and pneumonia
        colnames(data) <- c("hospital", "state", "heart attack", "heart failure", "pneumonia")
        # Check that state and outcome are valid
        
        if (!state %in% data[ ,"state"])
                { stop('invalid state')
        } else if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) 
                {stop('invalid outcome')
        } else if (is.numeric(rank)){
                
                state_logical <- which(data[, "state"] == state)
                
                df_state <- data[state_logical, ]    # subset data for having state
       
                df_state[,eval(outcome)] <- as.numeric(df_state[, eval(outcome)])
        
                df_state <- df_state[order(df_state[, eval(outcome)], df_state[, "hospital"]), ]

                output  <- df_state[,"hospital"][rank]
               
                
        }
          else if (!is.numeric(rank)){
                if (rank == "best") {
                
                output<- best(state,outcome)
        
        }       else if (rank == "worst") {
                
                state_logical <- which(data[, "state"]==state)
                
                df_state <-data[state_logical,]
                
                df_state[, eval(outcome)] <-as.numeric(df_state[,eval(outcome)]) #turn the outcome value into numeric
               
                df_state <- df_state[order(df_state[,eval(outcome)], df_state[,"hospital"], decreasing =T),] #order the hospital: if some hospital have same values, use alphabetically order
                
                output <- df_state[,"hospital"][1] #only return the first one
        }
                else 
                {stop("invalid rank")
        }
                
        
                }
        
return(output)     
}
        
rankhospital("NC", "heart attack", "worst")

```

** Reference**

Part of the code refers from the following links:

![]https://rpubs.com/fhlgood/rw3a3

![]https://xmuxiaomo.github.io/2015/06/19/R-Programming-Assignment-3/

![]https://github.com/DanieleP/PA3-tutorial


