---
title: "Class Note: Exploratory Analysis"
author: "Andrew Abisha Hu"
date: "8/19/2018"
output: html_document
---

# **Table of content**

* **Week1**

        * Basic Plotting system
        
* **Week2**   

        * Lattice Plotting system
        * Intro to ggplot
        
* **Week3**    

        * Hierarchical Clustering
        * K-means Clustering
        * Principals Componenets Analysis
        
* **Week4**
        
        * Case Study
        
---
 


# Week 1
        
        
### Simple base plot
```{r setup, include=FALSE}
library(datasets)
data(cars)
with(cars,plot(speed,dist))
```

### Lattice System
```{r cars}
library(lattice)
state<- data.frame(state.x77, region=state.region)
xyplot(Life.Exp~ Income | region, data=state, layout = c(4,1))
```


### ggplot qplot
```{r pressure, echo=FALSE}
library(ggplot2)
data(mpg)
qplot(displ,hwy,data=mpg)
```

## Basic Plotting System in R
```{r}
#Basic Plotting System in R
hist(airquality$Ozone)
with(airquality, plot(Wind, Ozone))

#Set global graphics parameters
#par()

#check default
par("lty")

#plot: make scatterplot
#lines: add lines
#points: add points
#text: add text labels
#title: add annotations outside the plot
#axis: add axis ticks/labels
with(airquality, plot(Wind, Ozone))
title(main= "Ozone")

#Base plot with subset

with(airquality, plot(Wind,Ozone, main="Ozone and Wind in NYC"))
with(subset(airquality,Month==5), points(Wind, Ozone, col="blue"))
with(subset(airquality,Month!=5), points(Wind, Ozone, col="red"))
legend("topright", pch=1 , col= c("blue","red"), legend=c("May", "Other Months"))

#add a regression line

with(airquality, plot(Wind,Ozone, main="Ozone and Wind",pch= 20))
model <- lm(Ozone~Wind, airquality)
abline(model, lwd=2)


######Base plot demons####

x<- rnorm(100)
hist(x)
y<- rnorm(100)
hist(y)
plot(x,y)

#margin of plots
par(mar= c(2,2,2,2))
plot(x,y)
par(mar= c(4,4,2,2))

#pch= different plotting symbols

plot(x,y, pch=20)
plot(x,y,pch=19)

#Add title and label
title("ScatterPlot")
text(-2,-2, "label")
legend("topleft",legend="Data" )

# Add regression line
fit <- lm(y~x)
abline(fit)

#Set the width of line
abline(fit, lwd=3, col="blue")

#plot two rows together
z<- rpois(100,2)
par(mfrow=c(2,1)) #two rows, one col
plot(x,y,pch=20)
plot(x,z,pch=19)

#rearrange the margin
par("mar")
par(mar= c(2,2,1,1))
plot(x,y,pch=20)
plot(x,z,pch=19)

#different rows of plots
par(mfrow=c(2,2))
plot(x,y)
plot(x,z)
plot(z,x)
plot(z,y)


#Loading points separately . Subsetting based on grouping variables
par(mfrow=c(1,1))
x<- rnorm(100)
y<- x + rnorm(100)
g <- gl(2,50)
g<- gl(2,50, labels = c("Male", "Female"))
str(g)

plot(x,y, type= "n") #meaning: don't put the data there yet
points(x[g=="Male"], y[g=="Male"], col= "green")
points(x[g=="Female"], y[g=="Female"], col= "blue")



```


---

# Week 2

## Lattice Plotting System: Intro
```{r Lattice Plotting}

############Lattice plotting system#########


#xyplot:


library(lattice)
library(datasets)

#Simple scatter
xyplot(Ozone~Wind, data=airquality)

#seperated by month

#convert month to factor
airquality <- transform(airquality, Month= factor(Month))
xyplot(Ozone~ Wind | Month, data=airquality,layout= c(5,1))


#Lattice Panel Functions##

#Generate random data
set.seed(10)
x<- rnorm(100)
f <- rep(0:1, each=50)
y <- x + f - f*x + rnorm(100, sd=0.5)
f <- factor(f, labels = c("Group1", "Group2"))

#plot with 2 panels
xyplot( y~x | f, layout= c(2,1))


#Custom panel function (structure)

xyplot(y~x | f, panel= function(x,y,...) {
        panel.xyplot(x,y,...)#first, call the default panel function for xyplot
        panel.abline(h = median(y), lty=2) #add line at median
        panel.lmline(x,y, col=2) #slr line
        
}) 
```



## ggplot2 Introduction:

### **Simple version: qplot**
```{r Simple qplot}
######  ggplot2 : qplot#####
library(ggplot2)
str(mpg)



#simple qplot
qplot(displ, hwy, data=mpg, color= drv)#assign different colors by drv

#adding a geom: smooth?
qplot(displ, hwy, data=mpg, geom= c("point", "smooth"))


#histogram
qplot(hwy, data=mpg, fill= drv)#seperate by drv


#Facets: subsets of data indicated by a factor
#var at right hand side determines the cols of the panels
#var at left: determine rows
qplot(displ, hwy, data=mpg, facets= .~drv)
qplot(hwy, data= mpg, facets= drv~.)

```

### Basic components of a ggplot

1. A df
2. aesthetic mappings: how data area mapped to color
3. geoms: geometric, e.g. points, lines, shapes
4. facets: for conditional plots (subsetting)
5. stats: statistical transformations like binning, quantiles, smoothing
6. scales: what scale an aesthetic map uses( e.g. male= red)
7. coordinate system


### Basic structure for ggplot
```{r }
#Base

#example
g<- ggplot(mpg, aes(displ,hwy))
print(g) #no layers yet

#Assign points
p<- g + geom_point() #save and print ggplot object
print(p)

g+ geom_point() #auto print plot object


#More layers: smooth 
g + geom_point() + geom_smooth(method= "lm")

#More layers: facet (seperate by a factor var)
g + geom_point() + facet_grid(. ~drv) +geom_smooth(method= "lm")
```

### Annotation:

Labels: xlab(), ylab(), labs(), ggtitle()
"geom" functions has options to modify

### Modifying Aesthetics:
```{r }
#Constant value of colors
g+ geom_point(color= "steelblue", size =4, alpha= 1/2)

#Color seperated by factor variable
g + geom_point(aes(color=drv), size=4, alpha= 1/2)
```
### Modifying labels:
```{r }
g + geom_point(aes(color = drv)) + labs(title = "MPG", x= "DISPL", y= "HWY")
```

### Customizing the Smooth:
```{r}
g + geom_point(aes(color=drv), size= 2, alpha=1/2)+ geom_smooth(size=4, linetype = 3, method= "lm", se=FALSE)
```

### Changing the Theme:
```{r}
g + geom_point(aes(color=drv)) + theme_bw(base_family = "Times")
```


### Notes about Axis Limits:
```{r}
testdat <- data.frame(x = 1:100, y=rnorm(100))
testdat[50,2] <-100  ##set outlier
plot(testdat$x, testdat$y, type= "l", ylim=c(-3,3))


g<- ggplot(testdat, aes(x=x, y=y))
g +geom_line()
```


### Plotting the variables: Factor + Continuous

So now, how do we plot the relationship between categorical and continuous variables?

Previously, we just seperate the plotting by factor variables(e.g. drive type, or BMI type)

But now we need to seperate them into a factor and a continuous type. Hence, we need to :

Transform continuous into a certain range of factors: using cut() functions
```{r}
#Not able to access the data so just take notes

#Calculate the deciles of the data
cutpoints <- quantile(df$var, seq(0,1, length= 4), na.rm=TRUE)

df$newvar <- cut(df$var, cutpoints)

#see levels
levels(df$newvar)
```

### **Summary**
```{r}
#Base
g<- ggplot(df, aes(x,y))

#Add layer
        g+
        geom_point(...) +
        facet_wrap(var1~ var2, nrow= 2, ncol=4) +
        geom_smooth(method="lm", se=FALSE, col="steelblue") +
        theme_bw(base_family= "Avenir", base_size = 10) +
        labs(title= "", x="", y="")
```

---

# Week 3

### Hierarchical Clustering
```{r}
#random sets
set.seed(1234)
par(mar= c(0,0,0,0))
x<- rnorm(12, mean= rep(1:3, each=4), sd=0.2)
y<- rnorm(12, mean=rep(c(1,2,1),each=4),sd=0.2)
plot(x,y,col="blue", pch=19, cex=2)
text(x+ 0.05, y+0.05, labels= as.character(1:12))


#Calculate the distance
dataFrame <- data.frame(x=x, y=y)
dist(dataFrame)

#clustering (create a dendrogram)
distxy<- dist(dataFrame)
hClustering <- hclust(distxy)
plot(hClustering)
```

### K-means Clustering
```{r}
#Random sets
set.seed(1234)
par(mar= c(0,0,0,0))
x<- rnorm(12, mean= rep(1:3, each=4), sd=0.2)
y<- rnorm(12, mean=rep(c(1,2,1),each=4),sd=0.2)
plot(x,y,col="blue", pch=19, cex=2)
text(x+ 0.05, y+0.05, labels= as.character(1:12))

#kmeans functions
kmeansObj <- kmeans(dataFrame, centers=3)
names(kmeansObj)
kmeansObj$cluster

#plot the data
par(mar=rep(0.2,4))
plot(x,y, col=kmeansObj$cluster, pch=19, cex=2)
points(kmeansObj$centers, col=1:3, pch=3, cex=3, lwd=3)

#alt: heatmap
set.seed(1234)
dataMatrix <- as.matrix(dataFrame)[sample(1:12),]
kmeansObj2 <- kmeans(dataMatrix, centers =3)
par(mfrow=c(1,2), mar= c(2,4,0.1,0.1))
image(t(dataMatrix)[,nrow(dataMatrix):1],yaxt= "n")
image(t(dataMatrix)[,order(kmeansObj$cluster)],yaxt= "n")

```

### Principals Componenets Analysis and Singular Value Decomposition (SVD)

SVD:
If X is a matrix with each var in a column and each obs in a row then the SVD is a "matrix decomposition"

X= UDV^T

where the columns of U are orthogonal (left singular vectors), the columns of V are orthogonal (right singular vector) and D is diagonal matrix (singular values)

```{r}
#Data frame
hh<- hclust(dist(dataMatrix))
dataMatrixOrd <- dataMatrix[hh$order,]
par(mfrow=c(1,3))

#COmponenets of the SVD- u and v

svd1 <- svd(scale(dataMatrixOrd))
par(mfrow=c(1,3))
image(t(dataMatrixOrd)[, nrow(dataMatrixOrd):1])

```

---

# Week 4

## Case Study: Samsung
```{r}
library(data.table)
sam<- load("/users/andrewhu/desktop/samsungData.rda")
names(sam)[1:12]


#average acceleration for each subject
par(mfrow=c(1,2) , mar=c(5,4,1,1))
sam<- transform(sam, activity= factor(activity))
sub1 <- subset(sam, subject==1)
plot(sub1[,1], col= sub1$activity, ylab= names(sub1)[1])
plot(sub1[,2], col= sub1$activity, ylab= names(sub1)[2])
legend("bottomright", legend= unique(sub1$activity), col= unique(sub1$activity), pch=1)


#Plotting max acceleration for the first subject

par(mfrow=c(1,2))
plot(sub1[,10], pch=19, col=sub1$activity, ylab = names(sub1)[10])
plot(sub1[,11],pch=19, col=sub1$activity, ylab=names(sub1)[11])


#SVD::

svd1 <- svd(scale(sub1[,-c(562,563)]))
par(mfrow = c(1,2))
plot(svd1$u[,1], col=sub1$activity, pch=19)
plot(svd1$u[,2], col=sub1$activity, pch=19)


#k-means 
kClust<- kmeans(sub1[,-c(562,563)], centers =6, nstart=100)
table(kClust$cluster, sub1$activity)

#k-means setting a nstart
kClust<- kmeans(sub1[,-c(562,563)], centers =6)
table(kClust$cluster, sub1$activity)

```


## Case Study: Air pollution
```{r }
#Reading table 1999
pm0<- read.table("/users/andrewhu/desktop/pm25_data/RD_501_88101_1999-0.txt", comment.char = "#", header= FALSE, sep ="|", na.strings= "")
dim(pm0)
head(pm0)

#read the first line for names
cnames<- readLines("/users/andrewhu/desktop/pm25_data/RD_501_88101_1999-0.txt",1)
cnames #not splitted yet.

#split by |
cnames<- strsplit(cnames, "|", fixed= TRUE)
cnames #check


#Assign names to the table
names(pm0) <- cnames[[1]]
head(pm0) #there are some names with blanks, which is not valid

#Making legible names
names(pm0) <- make.names(cnames[[1]])
head(pm0) #now the blanks are replaced with .


#basic checking----

x0 <- pm0$Sample.Value
class(x0)
str(x0)
summary(x0)

#checking missing values
mean(is.na(x0)) #11% are missing


#pm1

#Reading the second table
pm1<- read.table("/users/andrewhu/desktop/pm25_data/RD_501_88101_2012-0.txt", comment.char = "#", header= FALSE, sep ="|", na.strings= "")

dim(pm1)

names(pm1)<- make.names(cnames[[1]])
head(pm1)

#Simply compare.
x1 <- pm1$Sample.Value
summary(x1) #Note: median 7.63, lower than 1999. the min is negative, kinda strange.
str(x1)
mean(is.na(x1)) #5% missing values, less than 1999

#boxplot
boxplot(x0,x1)
boxplot(log10(x0),log10(x1)) 

#Examing negative values
negative <- x1<0
str(negative)
sum(negative, na.rm=TRUE)#total NA
mean(negative, na.rm=TRUE) #the proportion of NA


#dates, examing and transforming
dates<- pm1$Date
str(dates)
dates<- as.Date(as.character(dates),"%Y%m%d")
str(dates)

hist(dates,"month")
hist(dates[negative], "month")#check which period has more negative values


#####   Check monitors for one location only:   NY####

site0 <-unique(subset(pm0, State.Code==36, c(County.Code, Site.ID)))

site1 <- unique(subset(pm1, State.Code==36, c(County.Code, Site.ID)))

head(site0)

#paste county code and site ID together
site0 <- paste(site0[,1], site0[,2],sep=".")
site1 <- paste(site1[,1], site1[,2],sep=".")
str(site0)
str(site1)

#Finding the intersect bewteen site0 and site1
both<- intersect(site0,site1)
both

#Also , finding obs in each monitors

#Create a var for pasting the county code and site ID
pm0$county.site <- with(pm0, paste(County.Code, Site.ID, sep="."))

pm1$county.site <- with(pm1, paste(County.Code, Site.ID, sep="."))

#Subset to just NYC
cnt0 <- subset(pm0, State.Code==36 & county.site %in% both)
cnt1 <- subset(pm1, State.Code==36 & county.site %in% both)

head(cnt0)


#Split by county.side
split(cnt0, cnt0$county.site) #not useful

#Counting the rows , spliting by county site
sapply(split(cnt0, cnt0$county.site), nrow)
sapply(split(cnt1, cnt1$county.site), nrow)

#Targeting on one monitor: 63.2008
pm0sub <-subset(pm0, State.Code==36 & County.Code==63 & Site.ID == 2008)

pm1sub <-subset(pm1, State.Code==36 & County.Code==63 & Site.ID == 2008)

dim(pm1sub)
head(pm1sub)
dim(pm0sub)

#Plot the dates and pm25 value
dates1 <- pm1sub$Date
dates1 <- as.Date(as.character(dates1),"%Y%m%d")
x1sub <- pm1sub$Sample.Value
str(dates1)
plot(dates1, x1sub)


dates0 <- pm0sub$Date
dates0 <- as.Date(as.character(dates0),"%Y%m%d")
x0sub <- pm0sub$Sample.Value
plot(dates0, x0sub)



#Panel plot
par(mfrow= c(1,2), mar= c(4,4,2,1))

plot(dates0,x0sub, pch=20)
abline(h=median(x0sub,na.rm=T))

plot(dates1,x1sub, pch=20)
abline(h=median(x1sub,na.rm=T))


#Put the two plots on the same range
range(x0sub,x1sub, na.rm=T)
rng<- range(x0sub,x1sub, na.rm=T)

par(mfrow= c(1,2))
plot(dates0,x0sub, pch=20, ylim=rng)
abline(h=median(x0sub, na.rm=T))


plot(dates1,x1sub, pch=20, ylim=rng)
abline(h=median(x1sub, na.rm=T))




##### Look at each individual states

###See if each individual states improve or not across the years###

#Take avg value by state : perfect timing for using tapply

mn0<- with(pm0, tapply(Sample.Value, State.Code, mean, na.rm=T))
summary(mn0)

mn1<- with(pm1, tapply(Sample.Value, State.Code, mean, na.rm=T))
summary(mn1)


#Create DF
d0<- data.frame(state= names(mn0), mean=mn0)
d1<- data.frame(state= names(mn1), mean=mn1)


mrg <- merge(d0,d1, by="state")
dim(mrg)
head(mrg)

#Plot the data####

par(mfrow=c(1,1))
with(mrg, plot(rep(1999,52), mrg[,2],xlim= c(1998,2013)))
with(mrg, points(rep(2012,52), mrg[,3],xlim= c(1998,2013)))
segments(rep(1999,52), mrg[,2],rep(2012,52), mrg[,3])
```