---
title: "Machine Learning Summary"
output: html_document
---

# Data Splitting

*Libraries needed: (data for example)*
```{r setup, include=FALSE}
library(caret); library(kernlab); data(spam); library(ISLR); library(ggplot2); library(gridExtra);data(Wage) ;library(Hmisc)
```


*Training and test set*
```{r simple training and test}
inTrain <- createDataPartition(y=spam$type, p=0.75, list=FALSE)
training <- spam[inTrain,]
testing <- spam[-inTrain,]
```

*Fit a Linear model and check the final model*
```{r pressure, echo=FALSE}
modelFit <- train(type ~.,data=training, method="glm") 
modelFit$finalModel
```

*Use model to predict the test set:*
```{r}
predictions <- predict(modelFit,newdata=testing)
```

*Confusion matrix: Checking the accuracy*
```{r}
confusionMatrix(predictions,testing$type) #use predictions to check the outcome in the test set
```

*K-fold:*
```{r}
folds <- createFolds(y=spam$type,k=10,
                             list=TRUE,returnTrain=TRUE)
sapply(folds,length)
```

*Feature plot: reveal the relationship between each variables*
```{r}
featurePlot(x=training[,c("age","education","jobclass")],
            y = training$wage,
            plot="pairs")
```

----

# Standardizing: Dealing with skewness

*Center and scale:*
```{r}
modelFit <- train(type ~.,data=training,
                  preProcess=c("center","scale"),method="glm")
```

*Box-Cox:*
```{r}
preObj <- preProcess(training[,-58],method=c("BoxCox"))
trainCapAveS <- predict(preObj,training[,-58]$capitalAve)
```


*Imputing data (dealing with missing values):*
```{r}
preObj <- preProcess(training[,-58],method="knnImpute")
capAve <- predict(preObj,training[,-58]$capAve)
```

---

# Building features:

*Dummies:*
```{r}
dummies <- dummyVars(wage ~ jobclass,data=training)
head(predict(dummies,newdata=training))
```

*Check near-zero-covariates*
```{r}
nsv <- nearZeroVar(training,saveMetrics=TRUE)
```

---

# PCA (Combining features)

```{r}
modelFit <- train(training$type ~ .,method="glm",preProcess="pca",data=training)
confusionMatrix(testing$type,predict(modelFit,testing))
```

---

# Predicting with linear regression:

*Fit a slr:*
```{r}
lm1 <- lm(eruptions ~ waiting,data=trainFaith)
summary(lm1)
#same us
lm <- train(eruptions~ waiting, method="lm",data=training)
summary(lm$finalModel)
```

*Prediction with interval:*
Use your model building with training set, to predict the test set
```{r}
pred1 <- predict(lm1,newdata=testFaith,interval="prediction")
```


