---
title: "Quiz 4"
output: html_document
---
#Q1

Load the vowel.train and vowel.test data sets:

```
library(ElemStatLearn)

data(vowel.train)

data(vowel.test)

```

Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit (1) a random forest predictor relating the factor variable y to the remaining variables and (2) a boosted predictor using the "gbm" method. Fit these both with the train() command in the caret package.

What are the accuracies for the two approaches on the test data set? What is the accuracy among the test set samples where the two methods agree?

Loading Libraries and data set
```{r }
library(ElemStatLearn)
library(caret)
library(dplyr)
data(vowel.train)
data(vowel.test)
```

Transforming data:
```{r}
set.seed(33833)
vowel.train <- mutate(vowel.train, y=factor(y))
vowel.test <- mutate(vowel.test, y=factor(y))
```


Fitting Model:
```{r}
#randomForest
rf_fit <- train(y ~.,method="rf",
              data=vowel.train, 
              trControl = trainControl(method="cv"),number=3)
#gbm model
gbm_fit <- train(y~., method="gbm", verbose=FALSE, data=vowel.train)
```

Making predictions:
```{r}
pred_rf <- predict(rf_fit, vowel.test)
pred_gbm <- predict(gbm_fit, vowel.test)
```

Checking accuracy using confusion matrix:
```{r}
confusionMatrix(pred_rf,vowel.test$y)
confusionMatrix(pred_gbm,vowel.test$y)
```

Combining models:
```{r}
#Create the combining data frame based on our rf and gbm models
predDF <- data.frame(pred_rf,pred_gbm,y=vowel.test$y)
#Use gam model to train with the combining data frame
combModFit <- train(y ~.,method="gam",data=predDF)
pred_combining <- predict(combModFit, vowel.test)
#Check the combining model's accuracy
confusionMatrix(pred_rf,pred_gbm)$overall[1]

#Similar calculation method
sum(pred_rf[predDF$pred_rf == predDF$pred_gbm] == 
        predDF$y[predDF$pred_rf == predDF$pred_gbm]) / 
    sum(predDF$pred_rf == predDF$pred_gbm)
```

#Q2

Load the Alzheimer's data using the following commands

```
library(caret)

library(gbm)

set.seed(3433)

library(AppliedPredictiveModeling)

data(AlzheimerDisease)

adData = data.frame(diagnosis,predictors)

inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]

training = adData[ inTrain,]

testing = adData[-inTrain,]

```

Set the seed to 62433 and predict diagnosis with all the other variables using a random forest ("rf"), boosted trees ("gbm") and linear discriminant analysis ("lda") model. Stack the predictions together using random forests ("rf"). What is the resulting accuracy on the test set? Is it better or worse than each of the individual predictions?


Loading datasets
```{r}
library(caret)

library(gbm)

set.seed(3433)

library(AppliedPredictiveModeling)

data(AlzheimerDisease)

adData = data.frame(diagnosis,predictors)

inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]

training = adData[ inTrain,]

testing = adData[-inTrain,]
```


Fitting model:
```{r}
#RF
rf_fit <- train(diagnosis~., method="rf", trControl = trainControl(method="cv"),number=3,data=training)
#gbm
gbm_fit <-train(diagnosis~., method="gbm",verbose=FALSE, data=training)
#lda
lda_fit <- train(diagnosis~., method="lda",data=training)
```

Predictions with each model:
```{r}
pred_rf <- predict(rf_fit, testing)
pred_gbm <- predict(gbm_fit, testing)
pred_lda<- predict(lda_fit,testing)
```

Checking Accuracy using confusionMatrix:
```{r}
confusionMatrix(pred_rf,testing$diagnosis)
confusionMatrix(pred_gbm,testing$diagnosis)
confusionMatrix(pred_lda,testing$diagnosis)
```


Combinign all the models:
```{r}
#Create the combining data frame
combDF <- data.frame(pred_rf,pred_gbm,pred_lda,diagnosis=testing$diagnosis)

#Use randomForest to create model, with the combining data set
rf_overall <- train(diagnosis~., method="rf",data=combDF)

#Use the combining model to predict test set
rf_overall_pred<- predict(rf_overall,testing)

#Check accuracy
confusionMatrix(rf_overall_pred,testing$diagnosis)
```

#Q3

Load the concrete data with the commands:
```
set.seed(3523)

library(AppliedPredictiveModeling)

data(concrete)

inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]

training = concrete[ inTrain,]

testing = concrete[-inTrain,]
```

Set the seed to 233 and fit a lasso model to predict Compressive Strength. Which variable is the last coefficient to be set to zero as the penalty increases? (Hint: it may be useful to look up ?plot.enet).


Loading:

```{r}
set.seed(3523)

library(AppliedPredictiveModeling)
library(elasticnet)
data(concrete)

inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]

training = concrete[ inTrain,]

testing = concrete[-inTrain,]
```


Fit model:
```{r}
set.seed(233)
lasso <- train(CompressiveStrength~., method="lasso",data=training)
plot.enet(lasso$finalModel,xvar=c("penalty"),use.color = T)
```

---

# Q5
Load the concrete data with the commands:
```
set.seed(3523)

library(AppliedPredictiveModeling)

data(concrete)

inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]

training = concrete[ inTrain,]

testing = concrete[-inTrain,]
```
Set the seed to 325 and fit a support vector machine using the e1071 package to predict Compressive Strength using the default settings. Predict on the testing set. What is the RMSE?

```{r}
set.seed(3523)
library(e1071)
library(AppliedPredictiveModeling)


data(concrete)

inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]

training = concrete[ inTrain,]

testing = concrete[-inTrain,]
```



```{r}
#svm model
fit <- svm(CompressiveStrength~., data=training)

pred<- predict(fit,testing)

RMSE(pred,testing$CompressiveStrength)

```
